Interval Probability for Fuzzy Quantum Theories

### 2016-07-13 prob.tex
prob.tex \subsection{Quantum Interval-valued Probability Measures} keep typing...

\begin{example}[Three-dimensional quantum 4-value interval-valued
probability measure] Given a three dimensional Hilbert space with
an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}$
defined as follow:
\begin{enumerate}
\item $\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$
and $\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess$.
\item When $\ket{\psi}$ is parallel to $\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
or $-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$, where
$0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$, we define $\bar{\mu}(\proj{\psi})=\likely$
and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\unlikely$.
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$.
\end{enumerate}
$\bar{\mu}$ cannot correspond to any quantum probability measure
because of the exact same reason as in example~\ref{Three-dimensional-three-value}.
Before we verify that $\bar{\mu}$ is a quantum interval-valued probability
measure, we need a notation for cross product. For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\begin{pmatrix}\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{pmatrix}$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{pmatrix}$, their cross product is defined as follow.
\[
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\begin{vmatrix}\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{vmatrix}\ket{0}+\begin{vmatrix}\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{vmatrix}\ket{1}+\begin{vmatrix}\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{vmatrix}\ket{2}\textrm{ ,}
\]
where $\begin{vmatrix}\alpha & \beta\\
\alpha' & \beta'
\end{vmatrix}=\alpha\beta'-\alpha'\beta$. Notice that we put the complex conjugate in the cross product, so
that the result of cross product is orthogonal to $\ket{\psi}$ and
$\ket{\phi}$. 
\begin{eqnarray*}
\ip{\psi\times\phi}{\psi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}=0\textrm{ ,}\\
\ip{\psi\times\phi}{\phi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}=0\textrm{ .}
\end{eqnarray*}
Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi\times\phi}$ need not be normalized as usual.

To verify $\bar{\mu}$ is a quantum interval-valued probability measure,
it is sufficient to check $\bar{\mu}\left(\proj{\psi_{1}}\tensor*[_{\mathscr{O}}]{+}{}\proj{\psi_{2}}\right)\subseteq\bar{\mu}\left(\proj{\psi_{1}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{2}}\right)$
for orthogonal $\ket{\psi_{1}}$ and $\ket{\psi_{2}}$. We can enumerate
all possible cases as follow. 
\begin{enumerate}
\item When $\ket{\psi_{1}}$ is $\ket{0}$, then $\ket{\psi_{2}}$ must
be parallel to $\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
for $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Then, 
\begin{eqnarray*}
 &  & \bar{\mu}\left(\proj{\psi_{1}}\tensor*[_{\mathscr{O}}]{+}{}\proj{\psi_{2}}\right)\\
 & = & \bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}\times\psi_{2}}\right)
\end{eqnarray*}
\end{enumerate}
\qed\end{example}

>>>>

\begin{example}{[}Three-dimensional quantum 4-value interval-valued
probability measure{]} Given a three dimensional Hilbert space with
an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}$
defined as follow:
\begin{enumerate}
\item $\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$
and $\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess$.
\item Let $\ket{\psi\left(\theta,\gamma\right)}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\phi\left(\theta,\gamma\right)}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$,
where $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$. we define
$\bar{\mu}(\proj{\psi\left(\theta,\gamma\right)})=\ket{\phi\left(\theta,\gamma\right)}=\likely$
and $\bar{\mu}(\mathbb{1}-\proj{\psi\left(\theta,\gamma\right)})=\bar{\mu}(\mathbb{1}-\proj{\phi\left(\theta,\gamma\right)})=\unlikely$.
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$.
\end{enumerate}
$\bar{\mu}$ cannot correspond to any quantum probability measure
because of the exact same reason as in example~\ref{Three-dimensional-three-value}.
Before we verify that $\bar{\mu}$ is a quantum interval-valued probability
measure, we need a notation for cross product. For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\begin{pmatrix}\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{pmatrix}$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{pmatrix}$, their cross product is defined as follow.
\[
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\begin{vmatrix}\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{vmatrix}\ket{0}+\begin{vmatrix}\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{vmatrix}\ket{1}+\begin{vmatrix}\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{vmatrix}\ket{2}\textrm{ ,}
\]
where $\begin{vmatrix}\alpha & \beta\\
\alpha' & \beta'
\end{vmatrix}=\alpha\beta'-\alpha'\beta$. Notice that we put the complex conjugate in the cross product, so
that the result of cross product is orthogonal to $\ket{\psi}$ and
$\ket{\phi}$. 
\begin{eqnarray*}
\ip{\psi\times\phi}{\psi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}=0\textrm{ ,}\\
\ip{\psi\times\phi}{\phi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}=0\textrm{ .}
\end{eqnarray*}
Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi\times\phi}$ need not be normalized as usual.

To verify $\bar{\mu}$ is a quantum interval-valued probability measure,
it is sufficient to check 
\begin{eqnarray*}
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{3}}\right) & \subseteq & \bar{\mu}\left(\proj{\psi_{1}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{2}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}}\right) & \subseteq & \bar{\mu}\left(\proj{\psi_{2}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{3}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{2}}\right) & \subseteq & \bar{\mu}\left(\proj{\psi_{3}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{1}}\right)
\end{eqnarray*}
 for every orthonormal basis $\mathcal{B}=\left\{ \ket{\psi_{1}},\ket{\psi_{2}},\ket{\psi_{3}}\right\} $.
We can enumerate all possible cases as follow. 
\begin{enumerate}
\item When $\ket{\psi_{1}}$ is $\ket{0}$, then $\ket{\psi_{2}}$ must
be parallel to $\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
for $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Then,
\[
\ket{\psi_{1}}\times\ket{\psi_{2}}=\begin{pmatrix}1\\
0\\
0
\end{pmatrix}\times\begin{pmatrix}0\\
\rme^{\rmi\gamma}\sin\theta\\
\cos\theta
\end{pmatrix}=\begin{pmatrix}0\\
-\cos\theta\\
\rme^{-\rmi\gamma}\sin\theta
\end{pmatrix}=\rme^{-\rmi\gamma}\begin{pmatrix}0\\
\rme^{\rmi\left(\pi+\gamma\right)}\sin\left(\frac{\pi}{2}-\theta\right)\\
\cos\left(\frac{\pi}{2}-\theta\right)
\end{pmatrix}
\]
WLOG, we can pick $\ket{\psi_{2}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
for $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$, and $\ket{\psi_{3}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
for $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$ 
\[
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}\times\psi_{2}}\right)=\likely=\imposs\tensor*[_{\mathscr{I}}]{+}{}\likely=\bar{\mu}\left(\proj{\psi_{1}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{2}}\right)
\]
\[
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}}\right)=\necess\subseteq\likely\tensor*[_{\mathscr{I}}]{+}{}\unlikely=\bar{\mu}\left(\proj{\psi_{2}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{1}\times\psi_{2}}\right)
\]
If ($0<\theta<\frac{\pi}{2}$ and $\pi\le\gamma<2\pi$) or $\theta=\frac{\pi}{2}$
, then 
\[
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}\times\psi_{2}}\right)=\unlikely=\imposs\tensor*[_{\mathscr{I}}]{+}{}\unlikely=\bar{\mu}\left(\proj{\psi_{1}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{2}}\right)
\]
Similarly, when $\ket{\psi_{1}}$ is $\ket{\ps}$, equation (\ref{eq:non-additive-vectors})
holds.
\item When $\proj{\psi_{1}}\tensor*[_{\mathscr{O}}]{+}{}\proj{\psi_{2}}=\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0}$,
then Similarly, when $\proj{\psi_{1}}\tensor*[_{\mathscr{O}}]{+}{}\proj{\psi_{2}}=\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps}$,
equation (\ref{eq:non-additive-vectors}) holds.
\end{enumerate}
\qed\end{example}
